---
title: "HW3 by AK4598"
author: "Aleya Khalifa"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem 1

We will explore the `instacart` dataset in this problem.

```{r, include = FALSE}
library(p8105.datasets)
data("instacart")
```

**write a short description of the dataset, noting the size and structure of the data, describing some key variables, and giving illustrative examples of observations.**

## Problem 1.1

How many aisles are there, and which aisles are the most items ordered from?

```{r}

```

## Problem 1.2

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r plot_items}

```

## Problem 1.3

Make a table showing the three most popular items in each of the aisles "baking ingredients", "dog food care", and "packaged vegetables fruits". Include the number of times each item is ordered in your table.

```{r table_aisles}

```

## Problem 1.4

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r table_orders}

```

# Problem 2

We will use BRFSS data (`brfss_smart2010`) in this problem.

```{r, include = FALSE}
data("brfss_smart2010")
```

## Clean the data:

-   Format the data to use appropriate variable names

-   Focus on the "Overall Health" topic

-   Include only responses from "Excellent" to "Poor"

-   Organize responses as a factor taking levels ordered from "Poor" to "Excellent"


## Problem 2.1

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}

```


## Problem 2.2

### Construct a dataset

Limit it to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 

```{r}

```


### Make a “spaghetti” plot 

Plot average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r plot_spaghetti}

```


## Problem 2.3

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r plot_panels}

```


# Problem 3

In this problem, we will use accelerometer data from 250 participants. The data come from the NHANES study.

Note: Variables `*MIMS` are the MIMS values for each minute of a 24-hour day starting at midnight.

```{r, include=FALSE}
accel <- read_csv("data/nhanes_accel.csv") %>%
  janitor::clean_names()

covar <- read_csv("data/nhanes_covar.csv", 
                  skip = 4) %>%
  janitor::clean_names()
```


## Problem 3.1

Load, tidy, merge, and otherwise organize the data sets. The final dataset includes all originally observed variables; exclude participants less than 21 years of age, and those with missing demographic data; and encode data with reasonable variable classes (i.e. not numeric, and using factors with the ordering of tables and plots in mind).

```{r}

```


## Problem 3.2 

Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category. Comment on these items.

```{r}

```


## Problem 3.3

Using your tidied dataset, aggregate across minutes to create a total activity variable for each participant. Plot these total activities (y-axis) against age (x-axis); your plot should compare men to women and have separate panels for each education level. Include a trend line or a smooth to illustrate differences. Comment on your plot.

```{r}

```


## Problem 3.4

Accelerometer data allows the inspection activity over the course of the day. Make a three-panel plot that shows the 24-hour activity time courses for each education level and use color to indicate sex. Describe in words any patterns or conclusions you can make based on this graph; including smooth trends may help identify differences.

```{r}

```


